---
title: "R example - Housing development in Iceland"
author: "Sigríður Ösp Sigurðardóttir"
date: "01/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Reading table and adjustments to data
```{r}

library(forcats)
library(ggplot2)
library(dplyr)

sos <- read.table("husnaedisverd_2017.csv", encoding = "UTF-8",  header=TRUE, sep=";")
sos$matssvaedi <- factor(sos$matssvaedi)
sos$undirmatssvaedi <- factor(sos$undirmatssvaedi)
```

```{r}
sos$kaupverd <- paste0(sos$kaupverd, "000")
sos$kaupverd <- as.numeric(sos$kaupverd)
```

#### New Variables created
```{r}
sos$fermetraverd <- c(sos$kaupverd/sos$birtm2)

sos$teg_eign_grof <- fct_recode(sos$teg_eign,  Sérbýli = "Einbýlishús", Íbúð = "Íbúðareign", Sérbýli = "Parhús", Sérbýli = "Raðhús")
```

#### New dataframe eignir created consisting of housing in three towns
```{r}
eignir <- dplyr::filter(sos, svfn%in%c("1400", "2000", "6000"))
```

#### Town codes replaced with names
```{r}
eignir$svfn <- factor(eignir$svfn)
eignir$svfn <- fct_recode(eignir$svfn, Hafnarfjörður="1400", Reykjanesbær = "2000", Akureyri = "6000")
```

#### Number of premisises in each town
```{r}
ggplot(data=eignir, aes(svfn)) + geom_bar() + xlab("Town") + ylab("Num premises") + ggtitle("Premisises per town")
```


#### Number of premises split into houses and apartments
```{r}
ggplot(data=eignir, aes(svfn, fill=teg_eign_grof)) + geom_bar() + xlab("Town") + ylab("Num premises") + labs(fill = "Type of housing") + ggtitle("Premisises per town")
```

#### Median and average found for number of premisises by price per square meter
###### Mean: Red, Average: Blue
```{r}
options(scipen=10000)
ggplot(data=eignir, aes(fermetraverd, fill=)) + geom_histogram(bins = 20) + xlab("Price per sqm") + ylab("Num premises") + geom_vline(aes(xintercept = mean(fermetraverd, na.rm=T)),col='red',size=1) + geom_vline(aes(xintercept = median(fermetraverd, na.rm=T)),col='blue',size=1) + ggtitle("Number of premises by price per sqm")
```

**Mean:**
`r mean(eignir$fermetraverd, na.rm=T)`

**Median:**
`r median(eignir$fermetraverd, na.rm=T)`

#### Number of premises by price
```{r}
ggplot(data=eignir, aes(kaupverd)) + geom_histogram(bins=20) + xlab("Price") + ylab("Num premises") + geom_vline(aes(xintercept = mean(kaupverd, na.rm=T)),col='red',size=1) + geom_vline(aes(xintercept = median(kaupverd, na.rm=T)),col='blue',size=1) + ggtitle("Number of premises by price")
```

**Mean:**
`r mean(eignir$kaupverd, na.rm=T)`

**Median:**
`r median(eignir$kaupverd, na.rm=T)`


#### Number of premises in each town by year built
```{r}
ggplot(data=eignir, aes(byggar, fill=svfn)) + geom_histogram(bins=20) + xlab("Year built") + ylab("Num premises") + labs(fill = "Town") + ggtitle("Premises by year built")

```


#### Number of houses by size
```{r}
ggplot(data=dplyr::filter(eignir, teg_eign_grof%in%c("Sérbýli")), aes(birtm2, fill=svfn)) + geom_histogram(bins=20) + xlab("Size (sqm)") + ylab("Num houses") + labs(fill = "Town") + ggtitle("Number of houses by size")
```


#### Number of apartments by price per square meter
```{r}
ggplot(data=dplyr::filter(eignir, teg_eign_grof%in%c("Íbúð")), aes(fermetraverd, fill=svfn)) + geom_histogram(bins=20) + xlab("price per sqm") + ylab("Num apartments") + labs(fill = "Town") + ggtitle("Number of apartments by price per square meter")
```


#### Comparison of size of premises and full price by town
```{r}
ggplot(data=eignir, aes(x=birtm2, y=kaupverd, color=svfn)) + geom_point() + xlab("Premises size") + ylab("Price") + labs(color = "Town") + facet_grid(. ~ svfn) + ggtitle("Comparison of size of premises and price")
```

#### Number of bathtubs and showers in premises by year built
###### This was more of an experiment and still has some errors
```{r}
nuverold <- sos[sos$byggar>1949,]

ggplot(nuverold %>% group_by(byggar) %>% 
         summarise(fjbkar=mean(fjbkar)), 
       aes(x = byggar, y = fjbkar)) +  geom_line(colour="blue", linetype="11", size=0.3, na.rm = TRUE) + geom_point(shape=4, colour="blue", size=3, na.rm = TRUE) + xlab("Year") + ylab("Bathtub average") + scale_x_continuous(breaks = seq(1950,2020,10)) + ggtitle("Number of bathtubs in premises by year built")

ggplot(nuverold %>% group_by(byggar) %>% 
         summarise(fjsturt=mean(fjsturt)), 
       aes(x = byggar, y = fjsturt)) +  geom_line(colour="blue", linetype="11", size=0.3, na.rm = TRUE) + geom_point(shape=4, colour="blue", size=3, na.rm = TRUE) + xlab("Year") + ylab("Shower average") + scale_x_continuous(breaks = seq(1950,2020,10)) + ggtitle("Number of showers in premises by year built")
       
```


#### Tables of the three towns by type of premises
```{r}
table(eignir$svfn, eignir$teg_eign_grof)
```

```{r}
prop.table(table(eignir$svfn, eignir$teg_eign_grof), margin=1)
```

#### tapply summaries
```{r}
tapply(eignir$fermetraverd, eignir$svfn, summary, na.rm=TRUE)

```

```{r}
serb <- dplyr::filter(eignir, teg_eign_grof%in%c("Sérbýli"))
tapply(eignir$birtm2, eignir$svfn, summary, na.rm=TRUE)

```











